---
title: "Conflict and household data in Banlgadesh"
output: html_document
date: "2024-11-12"
---

```{r libraries, include=FALSE}
library("plyr")                     # Load plyr package
library("dplyr")                    # Manipulate data
library("haven")                    # Load dta files
library("sf")                       # Spatial features
library("raster")                   # Raster data (but also shapefiles)
```


## 1. MATCH ACLED DATA WITH A BGD SHAPEFILE AND BIHS ADMIN CODES ##

## Load administrative divisions of households from BIHS ##

Note: BIHS does not have GPS coordinates of households, but offers codes for the following administrative levels: division, district, upazila, and union (rural councils).

I will match conflict data from ACLED to the lowest administrative level 4 in rural Bangladesh (unions). 

```{r bihs}
# Set directory
setwd("C:/Users/idabr/OneDrive - Oxford Policy Management Limited/DEEP Conflict Climate Bangladesh/Data/BIHS data")

# Import data file with administrative codes in Bangladesh
BIHS_admin <- read_dta("BIHSlocations.dta")

```

##  Load shapefile for Bangladesh ##

Subnational boundaries at adminsitrative level 4 (unions) coming from https://data.humdata.org/dataset/cod-ab-bgd

```{r bgd shapefile admin 4}
# Set working directory 
setwd("C:/Users/idabr/OneDrive - Oxford Policy Management Limited/DEEP Conflict Climate Bangladesh/Data/Administrative boundary shapefiles/bgd_adm_bbs_20201113_shp/bgd_adm_bbs_20201113_SHP")

# Load the shapefile
bgd_admin4 <- read_sf("bgd_admbnda_adm4_bbs_20201113.shp")

```

## Load ACLED data on conflict ##

Import the excel spreadsheet with conflict data from ACLED. Then match the geographical coordinates of each conflict event to the union from the Bangladesh shapefile 

Each conflict event is geo-located with longitude and latitude

```{r acled}
# Set working directory
setwd("C:/Users/idabr/OneDrive - Oxford Policy Management Limited/DEEP Conflict Climate Bangladesh/Data/ACLED")

# Import the ACLED data 
acled_data <- read.csv("Bangladesh_ACLED.csv")

# Move longitude before latitude
acled_data <- acled_data %>% relocate(latitude, .after = longitude)

# Turn ACLED data frame into spatial points data frame
acled_data_sf <- st_as_sf(acled_data, coords = c(23:24))


```

## Spatial join between ACLED data and administrative level 4 in Bangladesh ##

```{r spatial join acled admin data}
# Set projection as the same for both objects - using the World Geodetic System 1984 CRS
st_crs(acled_data_sf) = 4326
st_crs(bgd_admin4) = 4326

# Have a look at the spatial objects
acled_data_sf
bgd_admin4

# Perform a spatial join between a spatial features data frame with ACLED data and Bangladesh shapefile. This function will check whether each spatial point (geographical coordinates) falls inside each polygon and return a data frame with successfully matched information from both spatial objects.

acled_geo <- st_join(acled_data_sf, bgd_admin4)

# Turn the spatial points data frame into a regular data frame 
acled_geo <- as.data.frame(acled_geo)

# Export as csv file - this csv file matches conflict events from ACLED with administrative level 4 codes from the shapefile from OCHA
write.csv(acled_geo, "ACLED_geo.csv")
```


# Match geo-referenced data from ACLED with BIHS ##

In the BIHS data, the admin 4 code has between 5-6 digits. Admin 4 code is a combination of codes for administrative levels 1, 2, 3 that need to be broken down to ensure compatibility. 

The format that I will save admin 4 codes in will follow from BIHS: 5-6 digits.

```{r}
# Create a new data frame with compatible codes
# Remove unnecessary columns

acled_geo1 <- acled_geo[ , !names(acled_geo) %in% 
    c("ADM4_REF","ADM4ALT1EN", "ADM4ALT2EN")]

# Rename the code identifiers for administrative levels - remove BD
acled_geo1$ADM4_PCODE <- substr(acled_geo1$ADM4_PCODE, 3, 10)
acled_geo1$ADM3_PCODE <- substr(acled_geo1$ADM3_PCODE, 3, 10)
acled_geo1$ADM2_PCODE <- substr(acled_geo1$ADM2_PCODE, 3, 10)
acled_geo1$ADM1_PCODE <- substr(acled_geo1$ADM1_PCODE, 3, 10)

# Check that codes for admin 1 (division) follow the same conventions
head(acled_geo1$ADM1_PCODE)
head(BIHS_admin$dvcode)

# Yes - they follow two-digit convention

# Remove the first 2 digits from the union code (those refer to admin level 1)
acled_geo1$ADM4_PCODE <- substr(acled_geo1$ADM4_PCODE, 3, 10)

# This is correct for most cases. However, whenever "0" is the first digit, we should delete it 

# Change to numeric - this will drop 0 at the front 
acled_geo1$ADM4_PCODE <- as.numeric(acled_geo1$ADM4_PCODE)

# Back to character 
acled_geo1$ADM4_PCODE <- as.character(acled_geo1$ADM4_PCODE)

# Rename columns in BIHS_admin before the join 
names(BIHS_admin)[names(BIHS_admin) == "uncode"] <- "ADM4_PCODE"

# Back to character for compatibility of the merge
acled_geo1$ADM4_PCODE <- as.character(acled_geo1$ADM4_PCODE)
BIHS_admin$ADM4_PCODE <- as.character(BIHS_admin$ADM4_PCODE)

# Left join - this data will assing all conflict events to the unions in which BIHS households are located
BIHS_2 <- left_join(BIHS_admin, acled_geo1)

# Note that some unions did not have a any conflict events, while others had multiple conflict events. 

# Have a look at unique values of admin 4 in both data frames
acled_p4 <- unique(acled_geo1$ADM4_PCODE)
BIHS_p4 <- unique(BIHS_admin$ADM4_PCODE)

# ACLED data covers 1358 unique unions while BIHS covers 275 unique unions

# Remove ugly columns
acled_geo1 <- acled_geo1[ , !names(acled_geo1) %in% 
    c("date","validTo", "validOn")]

# Export ACLED data as a csv file where the "ADM4_PCODE" column corresponds to the "uncode" column from the BIHS data. 
write.csv(acled_geo1, "ACLED_geo1.csv")


```
