---
title: "Drought data Bangladesh"
output: html_document
date: "2024-11-12"
---

```{r libraries}
library("ncdf4")                    # package for netcdf manipulation
library("raster")                   # package for raster manipulation
library("SPEI")                     # SPEI construction
library("vctrs")                    # Vector operations
library("leaflet")                  # Interactive maps
library("htmlwidgets")              # HTML widgets
library("RColorBrewer")             # Nice colours
```

## 2. DROUGHT DATA ##

The aim is to construct SPEI (standardised precipitation evapo-transpiration index) for Bangladesh. SPEI is a multi-scalar index indicating drought conditions. It is calculated as the difference between precipitation and PET (potential evapotranspiration). Climate data is taken from the CRU (Climatic Research Unit) Version 4 gridded dataset and contains monthly values of PET and precipitation on a 0.5 degrees regular grid over the period 1981-2020. 

# PET (potential evapo-transpiration) data #

Load the PET data (potential evapo-transpiration) merged netCDF file. This data has been merged using cdo climate operators package in Linux and has monthly values of PET for the years 1981-2020 on regular 0.5 degree grid. 

Note: Since I already processed the same drought data for the Nigeria paper, I will use the same precipitation and PET data for efficiency.

```{r Load netCDF files, include=FALSE}
# Set the working directory to netCDF files 
setwd("C:/Users/idabr/OneDrive - Oxford Policy Management Limited/DEEP Multiple Crises Poverty Nigeria/Data/CRU Version 4 Climate Data/PET (Potential evapo-transpiration)/Raw data")

# Open the netCDF file. This has already been merged and covers the period 1981-2020. It contains 480 time steps (12 months across 40 years)
pet_data <- nc_open("merged_pet.nc", write=FALSE, readunlim=TRUE, verbose=FALSE, 
 	auto_GMT=TRUE, suppress_dimvals=FALSE, return_on_error=FALSE )

# Data has three dimensions: lon, lat, and time. The variable of interest is "pet" 

# Extract variables - three dimensions 
lon <- ncvar_get(pet_data, "lon")                # longitude
lat <- ncvar_get(pet_data, "lat", verbose = F)   # latitude
t <- ncvar_get(pet_data, "time")                 # time

# Inspect the units that are used in the time dimension
tunits <- ncatt_get(pet_data,"time","units")
tunits

# Store the dimensions of the time variable 
nt <- dim(t)
nt

# 480 time units 

# Look at the first few entries from the longitude variable
head(lat) # 360 values (1st:-89.75) with 0.5 unit spacing
head(lon) # 720 values(1st: -179.75) with 0.5 unit spacing
head(t)

# Extract the variable of interest (potential evapo-transpiration)
pet.array <- ncvar_get(pet_data, "pet") # store the data in a 3-dimensional array

# Checking the dimensions of the array
dim(pet.array)

# [1] 720 360 480
# 720 longitudes, 360 latitudes, and 480 time units (12 months across 40 years)

# See what fill value was used for missing data
fillvalue <- ncatt_get(pet_data, "pet", "_FillValue")
fillvalue

#[1] 9.96921e+36

# Replace missing values with the usual "NA"
pet.array[pet.array == fillvalue$value] <- NA

# Note: the array looks like it has loads of missing values. Check?
head(pet.array)

# Clear all
#rm(list = ls())

# Close the netCDF file
nc_close(pet_data)

```

Understand the time variable

Note: the way the time variable is coded in this dataset: each value is a number of days since 1900-1-1 in chronological order.

```{r Process time variable, include=FALSE}

# Convert time -- split the time units string into fields
tustr <- strsplit(tunits$value, " ")
tdstr <- strsplit(unlist(tustr)[3], "-")
tmonth <- as.numeric(unlist(tdstr)[2])
tday <- as.numeric(unlist(tdstr)[3])
tyear <- as.numeric(unlist(tdstr)[1])

#chron(time,origin=c(tmonth, tday, tyear))# note: this function does not seem to work with non-numeric or non-character values 


# Check the number of non-NA values
length(na.omit(as.vector(pet.array[,,1])))

# [1] 66501
```


Convert PET into a data frame - name columns according to time steps (monthly data covering 1981-2020). 

```{r Turn PET into a data frame}
# Create a matrix of lon-lat pairs 
lonlat <- as.matrix(expand.grid(lon,lat))
dim(lonlat)

# Make a vector with values for PET
pet_vec <- as.vector(pet.array)
length(pet_vec)

# reshape the vector into a matrix
pet_mat <- matrix(pet_vec, nrow=720*360, ncol=nt)
dim(pet_mat)

# Inspect the head of the matrix (excluding missing values)
head(na.omit(pet_mat))

# Create a dataframe using the lon-lat matrix 
pet_df <- data.frame(cbind(lonlat,pet_mat))

# Assign names according to the original time dimension of the data (days since 1900-1-1)
names(pet_df) <- c("lon","lat", t)
# options(width=96)
head(na.omit(pet_df, 20))

# Now we have a data frame where each column is a point in time (12 months over 40 years) - need to break those down into month and year 

# Create a matrix of month-year combinations 

months <- 1:12
years <- 1981:2020
month_names <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
month_year <- as.matrix(expand.grid(months,years))
dim(month_year)

# Make a list with all combinations of months and years 
timeref <- list()

# Need to start the loop with the year then month
for (year in years) {
  timeref[[length(timeref)+1]] <- paste(year, month_names)
}


# Turn a list into a vector
timeref_vector <- unlist(timeref)

# Assign month-year combinations as column names for identification
names(pet_df) <- c("lon", "lat", timeref_vector)


```

Leave only points that are relevant for Bangladesh as defined by the shapefile. 

```{r shapefile}
# Set working directory 
setwd("C:/Users/idabr/OneDrive - Oxford Policy Management Limited/DEEP Conflict Climate Bangladesh/Data/Administrative boundary shapefiles/bgd_adm_bbs_20201113_shp/bgd_adm_bbs_20201113_SHP")

# Load the shapefile
bgd_admin4 <- read_sf("bgd_admbnda_adm4_bbs_20201113.shp")

# Check the extent 
extent(bgd_admin4)
```


Keep only values that fall within the extent of Bangladesh. Extent is given below 

92.75 > X (longitude) > 87.75
26.75 > Y (latitude) > 20.25


class      : Extent 
xmin       : 88.00863 
xmax       : 92.68031 
ymin       : 20.59061 
ymax       : 26.63451 

Note: in the data set negative latitudes represent the southern hemisphere and negative longitudes represent the western hemisphere.

```{r extent}
# Manually adjust extent 
pet_df = pet_df[pet_df$lon >= 87.75 & pet_df$lon <= 92.75,]
pet_df = pet_df[pet_df$lat >= 20.25 & pet_df$lat <= 26.75,]

# Make a new data frame in case something goes wrong 
pet_geo <- pet_df

# Make row names nicer
rownames(pet_geo) <- 1:nrow(pet_geo)

# Remove massive data files to free up space
rm(pet_df, pet_data, pet_vec, pet.array)

```


# Change the units of PET data. The documentation for CRU Version 4 climate data says PET is expressed in mm/day while Precipitation Rate is expressed in mm/month. Going to convert PET into monthly values to match precipitation.

```{r Unit conversion PET}

# Change daily values to monthly - multiply by 30.
pet_geo[,3:482] <- pet_geo[,3:482]*30

```


##  Precipitation data ##

# Load and process raw netCDF file with merged precipitation data (monthly for the period 1981-2020).

```{r Open precipitation data, include=FALSE}
# Set the working directory to .dat files as an experiment 
setwd("C:/Users/idabr/OneDrive - Oxford Policy Management Limited/DEEP Multiple Crises Poverty Nigeria/Data/CRU Version 4 Climate Data/PRE (Precipitation)/Raw data")

# Open the netCDF file. This has already been merged and covers the period 1981-2020. It contains 480 time steps (12 months across 40 years)
pre_data <- nc_open("merged_pre.nc", write=FALSE, readunlim=TRUE, verbose=FALSE, 
 	auto_GMT=TRUE, suppress_dimvals=FALSE, return_on_error=FALSE )

# Data has three dimensions: lon, lat, and time. The variable of interest is "pet" 

# Extract variables - three dimensions 
lon <- ncvar_get(pre_data, "lon")                # longitude
lat <- ncvar_get(pre_data, "lat", verbose = F)   # latitude
t <- ncvar_get(pre_data, "time")                 # time

# Inspect the units that are used in the time dimension
tunits <- ncatt_get(pre_data,"time","units")
tunits

# Store the dimensions of the time variable 
nt <- dim(t)
nt

# 480 time units 

# Look at the first few entries from the longitude variable
head(lat) # 360 values (1st:-89.75) with 0.5 unit spacing
head(lon) # 720 values(1st: -179.75) with 0.5 unit spacing
head(t)

# Extract the variable of interest (potential evapo-transpiration)
pre.array <- ncvar_get(pre_data, "pre") # store the data in a 3-dimensional array

# Checking the dimensions of the array
dim(pre.array)

# [1] 720 360 480
# 720 longitudes, 360 latitudes, and 480 time units (12 months across 40 years)

# See what fill value was used for missing data
fillvalue <- ncatt_get(pre_data, "pre", "_FillValue")
fillvalue

#[1] 9.96921e+36

# Make some space by removing the massive pet_data
rm(pet_data)


# Need more space - keep just objects for Nigeria
rm(pet_mat, pet_df)
rm(pet.array, pet_vec)

# Replace missing values with the usual "NA"
pre.array[pre.array == fillvalue$value] <- NA

# Note: the array looks like it has loads of missing values. Check?
head(pre.array)

# Clear all
#rm(list = ls())

# Close the netCDF file
nc_close(pre_data)
```


Turn the precipitation netCDF file into a data frame. Assign intuitive markers for time (months from 1981-2020).

```{r Precipitation data frame, include=FALSE}

# Make a vector with values for PRE
pre_vec <- as.vector(pre.array)
length(pre_vec)

# reshape the vector into a matrix
pre_mat <- matrix(pre_vec, nrow=720*360, ncol=nt)
dim(pre_mat)

# Create a dataframe using the lon-lat matrix 
pre_df <- data.frame(cbind(lonlat,pre_mat))

# Assign names according to the original time dimension of the data (days since 1900-1-1)
names(pre_df) <- c("lon","lat", t)
# options(width=96)
head(na.omit(pre_df, 20))

# Now we have a data frame where each column is a point in time (12 months over 40 years) - need to break those down into month and year 

# Assign month-year combinations as column names for identification
names(pre_df) <- c("lon", "lat", timeref_vector)
```


Restrict the sample to data points in Bangladesh - using coordinates from the PET data frame which has been spatially joined with the Bangladesh shapefile 

```{r Bangladesh precipitation}
# Make a data frmae with just 0.5 degree coordinates that fall within Bangladesh
BG_coords <- pet_geo[,c("lon", "lat")]

# Left join with the precipitation data frame
pre_geo <- left_join(BG_coords, pre_df)

# Remove large unnecessary objects to free up space 
rm(pre_data, pre_df, pre_mat, pre_vec, pre.array)
```

##  SPEI construction ##

Construct SPEI using the SPEI package in R. As a first step I need a time series of the water balance (precipitation minus potential evapotranspiration). The two data frames with these variables need to have exactly the same dimensions to subtract them from each other. 

```{r Calculate water balance}
# Create a data frame for PET within Bangladesh that has only columns relating to PET (and not info from the shapefile)
pet_bgd <- pet_geo[,1:482]

# Subtract PET from precipitation
water_balance <- pre_geo[,3:482] - pet_bgd[,3:482]

# Append the longitude and latitude
water_balance['lon'] <- pre_geo$lon
water_balance['lat'] <- pre_geo$lat
  
# Move longitude and latitude to the front
water_balance <- water_balance %>% 
  relocate(lat)

water_balance <- water_balance %>% 
  relocate(lon)

```


Construct SPEI. Data needs to be in the following format: Water balance needs to be a column with monthly values of a time series. 

SPEI package parameters:

1) Scale parameter controls the influence of past values. For example, selecting 12 will take into account 12 previous months.
2) Distribution parameter decides what kind of distribution the data should be fit to. For some reason, log-logistic returns all values while Gamma comes up with NA. I will use log-logistic below.
3) Reference period: default will be taking into account the whole reference period of the data. In my case that is 1981-2020.

Note: do we want drought measured on different time scales? 1,2,3,4,5 year cumulative drought? 

Brunckhorst (2020) uses a 12-month SPI. Take the value of December for each year.

Calculate 12-month Dec SPEI for all years 2010-2020. In order not to spend 3 years on this, it would be useful to have a loop that goes through these individual years, calcultes SPEI and saves the result as a column in a data frame with lon-lat and X-Y combinations 

```{r turn water balance into ts}
# First we need to turn the water balance data frame into a time series 

# Turn lon lat into characters and add column where both are displayed
BG_coords$lon <- as.character(BG_coords$lon)
BG_coords$lat <- as.character(BG_coords$lat)
BG_coords$lonlat <- paste(BG_coords$lon, BG_coords$lat, sep = " ")

# Add the column for lon-lat
water_balance$lonlat <- BG_coords$lonlat

# Delet individual lon-lat measurements
water_balance <- water_balance[,3:483]

# Move lon-lat to the front
water_balance <- water_balance %>% 
  relocate(lonlat)

# Transpose to make time-series a column
water_balance_long <- as.data.frame(t(water_balance))

# Make column names lon-lat
names(water_balance_long) <- water_balance_long[1,]

# Make columns numeric values
water_balance_long <- sapply(water_balance_long, as.numeric)

# Remove lon-lat
water_balance_long <- water_balance_long[-1,]

# Declare water balance a time series, specifying the start and end point

water_balance_long <- ts(water_balance_long, start=c(1981,1), end=c(2020,12), frequency=12)

```

SPEI loop

```{r}
# Store the sequence of years in a vector
s_years <- 2010:2020

# Create a sequence of months
all_months <- vec_rep(month_names, 40)

# Create a sequence of years
all_years <- rep(years,each=12)


# Create a list to store fitted values
fitted_list <- list()

# Loop time

for (i in s_years) {
spei_list <- spei(water_balance_long, 12, kernel = list(type = 'rectangular', shift = 0),  # Calculate 12-month SPEI 
distribution = 'log-Logistic', fit = 'ub-pwm', na.rm = TRUE,
ref.start=NULL, ref.end=c(i,12), x=FALSE, params=NULL)
fitted <- spei_list$fitted 
fitted <- as.data.frame(fitted) 
fitted$month <- all_months
fitted$year <- all_years
data <- fitted %>% filter(month =="Dec" & year==i)
data <- data[,1:154] # number of grid cells
data <- as.data.frame(t(data))
data$lon <- BG_coords$lon   # Assign lon and lat
data$lat <- BG_coords$lat
fitted_list[[i-2009]] <- data

}


# Prepare a data frame
spei_2010_2020 <- pet_geo[,1:2]

spei_2010_2020 <- fitted_list[[1]]
colnames(spei_2010_2020)[which(names(spei_2010_2020) == "V1")] <- "SPEI_2010"


values <- list()
for (i in s_years) {
  values[[i-2009]] <- fitted_list[[i-2009]]$V1
}

spei_2010_2020$SPEI_2011 <- values[[2]]
spei_2010_2020$SPEI_2012 <- values[[3]]
spei_2010_2020$SPEI_2013 <- values[[4]]
spei_2010_2020$SPEI_2014 <- values[[5]]
spei_2010_2020$SPEI_2015 <- values[[6]]
spei_2010_2020$SPEI_2016 <- values[[7]]
spei_2010_2020$SPEI_2017 <- values[[8]]
spei_2010_2020$SPEI_2018 <- values[[9]]
spei_2010_2020$SPEI_2019 <- values[[10]]
spei_2010_2020$SPEI_2020 <- values[[11]]


# Rename rows
rownames(spei_2010_2020) <- 1:nrow(spei_2010_2020)

# Move longitude and latitude to the front
spei_2010_2020 <- spei_2010_2020 %>% 
  relocate(lat)

spei_2010_2020 <- spei_2010_2020 %>% 
  relocate(lon)

# Remove grid cells with NA values
spei_2010_2020 <- spei_2010_2020[!is.na(spei_2010_2020$SPEI_2010),]

# Make row names nices
rownames(spei_2010_2020) <- 1:nrow(spei_2010_2020)

# Export this as a csv file
write.csv(spei_2010_2020, "SPEI_2010_20_BGD.csv")
```


## Visualise the data a little to make sure it works ##

Take SPEI values for 2010.

```{r}

# Test for 2010
spei_2010 <- spei_2010_2020[,1:3]

# Turn values from 2010 into a raster
r_drought <- rasterFromXYZ(spei_2010)
plot(r_drought) # quick check that raster is populated with values

# Define colour palette for the map - red indicates lower values of SPEI and yellow higher values. 
pal1 <- colorNumeric(c("#B10026", "#E31A1C", "#FFFFCC", "#FC4E2A", "#FD8D3C", "#FEB24C", "#FED976", "#FFEDA0"), values(r_drought),
  na.color = "transparent")

# Assign a CRS to the raster
crs(r_drought) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" 

# Create an interacive map showing SPEI values for 2010 
map5 <- leaflet() %>%
 addProviderTiles(providers$CartoDB.Positron) %>%
  addRasterImage(r_drought, colors = pal1, opacity=0.7, project = FALSE)%>%
  addLegend(pal = pal1, values = values(r_drought), title = "SPEI") 

map5
```

