---
title: "Climate visualisation Bangladesh"
output: html_document
date: "2024-11-14"
---

```{r library}
library("plyr")                     # Load plyr package
library("dplyr")                    # Data manipulation
library("readr")                    # Load and export csv files
library("raster")                   # Raster data
library("sf")                       # Spatial features
library("tidyverse")                # Data manipulation
library("leaflet")                  # Interactive maps
library("htmlwidgets")              # HTML widgets
library("RColorBrewer")             # Nice colours
library("ggplot2")                  # package for plotting
library("exactextractr")            # Zonal statistics using geo-spatial data
library("reshape")                  # Melt long
library("scales")                   # Nice scales in graphs
```


## Visualise floods ##

Turn layers of each flood event in Bangladesh between 2010-2018 into a mosaic, i.e. layer all floods on top of each other. Map all flooded areas in Bangladesh between 2010-2018

```{r flood vis}
# Import base raster layer for Bangladesh from WorldPop - this details the boundaires of the country 
base.layer <- raster('C:/Users/idabr/OneDrive - Oxford Policy Management Limited/DEEP Conflict Climate Bangladesh/Data/Base raster/bgd_level0_100m_2000_2020.tif')

# Set the working directory to TIF files
setwd("C:/Users/idabr/OneDrive - Oxford Policy Management Limited/DEEP Conflict Climate Bangladesh/Data/Global Flood Database")

# Create a list of TIF files using the DFO pattern and save as rasters
DFO_files <- list.files(pattern = "*DFO",
                           full.names = TRUE) %>%
  lapply(raster)

# Extract only the layer we are interested in (flooded) - for all these files. Replace all elements in the list 
for (i in 1:39) {
 DFO_files[[i]] <- subset(DFO_files[[i]], "flooded")
  
}

# Set extent of each raster layer to that of the base layer - this will give the max and min coordinates that bound Bangladesh

for (i in 1:39) {
DFO_files[[i]] <- crop(DFO_files[[i]], base.layer)  

}

# Create an initial mosaic with the first two rasters
megaraster <- mosaic(DFO_files[[1]], DFO_files[[2]], fun="max")

# Turn into a mosaic 
for (i in 3:39) {

megaraster <- mosaic(megaraster, DFO_files[[i]], fun="max")

}


# Mask to get only the shape of Bangladesh

# Need the same spatial resolution and extent between the base raster layer and flood layers
base.layer <- resample(base.layer, megaraster, method = "ngb") # resampling nase raster to the spatial resolution of the flood layers
megaraster <- crop(megaraster, base.layer) # setting extent of flood layer same as base layer
megaraster <- mask(x = megaraster, mask = base.layer) # masking to the shape of Bangladesh from the base layer 

# Reclassify 0s to missing values so they are not displayed on the map
y <- reclassify(megaraster, cbind(-Inf, 0, NA))

# Check that there are no missings
summary(y)

# Quick preview
plot(y, axes=F, box=F, legend=F)

# Create an interactive map
# Set colour palette 
pal <- colorNumeric(c("#FFFFCC", "#0C2C84" ,"#41B6C4"), values(y),
  na.color = "transparent")

map2 <- leaflet() %>%
 addProviderTiles(providers$CartoDB.Positron) %>%
  addRasterImage(y, colors = pal) %>%
  addLegend(pal = pal, values = values(y),
    title = "Floods")

map2

# Save map as widget
saveWidget(map2, file="BGD_floods.html")

```

## Add droughts to the visualisation ##

Map the worst drought in Bangladesh between 2010-2020

```{r vis drought}
# Import the data frame with SPEI values
setwd("C:/Users/idabr/OneDrive - Oxford Policy Management Limited/DEEP Conflict Climate Bangladesh/Code")

# Make a data frame with just geographical coordinates, regular 0.5 degrees grid and SPEI values 
drought_201020 <- read.csv("SPEI_2010_20_BGD.csv")

# Add a column that has the lowest value of SPEI between 2010-2020
drought_201020$minSPEI <- apply(drought_201020[,4:14],1,min)

# Replace all values above -1.5 with NA (we are only interested in drought defined as SPEI<-1.5)
drought_201020$minSPEI[drought_201020$minSPEI > -1.5] <- NA

# Let's also have a version for SPEI<0
drought_201020$minSPEIzero <- apply(drought_201020[,4:14],1,min)

# Replace all values above 0 with NA (we are only interested in drought defined as SPEI<0)
drought_201020$minSPEIzero[drought_201020$minSPEI > 0] <- NA

# Create a data frame with just the variable of interest and grid coordinates 

ex_drought <- drought_201020[, c("lon", "lat", "minSPEI")]
spei_zero_drought <- drought_201020[, c("lon", "lat", "minSPEIzero")]

# Remove missing values
ex_drought <- ex_drought[complete.cases(ex_drought), ]
spei_zero_drought <- spei_zero_drought[complete.cases(spei_zero_drought), ]

# Create raster
r_drought <- rasterFromXYZ(ex_drought)
r_drought_zero <- rasterFromXYZ(spei_zero_drought)

# Mask to the shape of Bangladesh
r_drought <- crop(r_drought, base.layer)  # set extent the same as base layer
base.layer <- resample(base.layer, r_drought, method = "ngb") # resample base layer to the same spatial resolution as drought raster (0.5 degrees)
r_drought <- mask(x = r_drought, base.layer) # mask to get just the shape of Bangladesh

# Simple plot to see how the raster came out

plot(r_drought, axes=F, box=F)

# Tag a coordinate system
crs(r_drought) <- sp::CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
crs(r_drought_zero)  <- sp::CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# Check colour palettes - want something red
display.brewer.pal(n = 8, name = 'YlOrRd')

# Hexadecimal color specification 
brewer.pal(n = 8, name = "YlOrRd")

# Set colour palette 
pal1 <- colorNumeric(c("#B10026", "#E31A1C", "#FFFFCC", "#FC4E2A", "#FD8D3C", "#FEB24C", "#FED976", "#FFEDA0"), values(r_drought),
  na.color = "transparent")

pal2 <- colorNumeric(c("#E31A1C", "#FFFFCC", "#FC4E2A", "#FD8D3C", "#FEB24C", "#FED976", "#FFEDA0"), values(r_drought_zero),
  na.color = "transparent")


# Now put it onto leaflet 
map3 <- leaflet() %>%
 addProviderTiles(providers$CartoDB.Positron) %>%
  addRasterImage(r_drought,  colors = pal1, project=FALSE)%>%
   addLegend(pal = pal1, values = values(r_drought), title = "SPEI")

map3

```

## Combine visualisations of flood and drought in Bangladesh ##


```{r}
# Create a colour pallete for drought values
pal3 <- colorNumeric(c("#E31A1C", "#FFFFCC", "#FC4E2A", "#FD8D3C", "#FEB24C", "#FED976", "#FFEDA0"), values(r_drought),
  na.color = "transparent")

# Make sure raster image with drought has the same CRS as floods
crs_flood <- crs(y)
crs(r_drought) <- crs_flood

map5 <- leaflet() %>%
 addProviderTiles(providers$CartoDB.Positron) %>%
  addRasterImage(r_drought,  colors = pal3, opacity=0.7, project = FALSE)%>%
  addLegend(pal = pal3, values = values(r_drought), title = "SPEI") %>%
  addRasterImage(y, colors = pal) 

map5

# Save map as widget
saveWidget(map5, file="BGD_climate_map.html")
```


# Make graphs with trends in droughts and floods over time #

For floods, we can simply count the number of flood events occuring each year. For drought, I will show the number of districts (there are 64 districts in Bangladesh in total) affected by drought each year. District is administrative level 2

New approach (Vidya's idea): for each year and each state, find the minimum value of SPEI and assign it to one of the 4 categories below so that we get non-overlapping categories in the stacked area graph.

These are the 4 categories of drought:
1.  'SPEI < -1.5'
2. '-1.5 < SPEI < -1'
3.  '-1 < SPEI < -0.5'
4. "-0.5 < SPEI < 0"

```{r}
# Import shapefile with administrative level 2 (district)
bgd_adm2 <- read_sf("C:/Users/idabr/OneDrive - Oxford Policy Management Limited/DEEP Conflict Climate Bangladesh/Data/Administrative boundary shapefiles/bgd_adm_bbs_20201113_shp/bgd_adm_bbs_20201113_SHP/bgd_admbnda_adm2_bbs_20201113.shp")

# Remove first column with "X"
drought_201020 <- drought_201020[,2:16]

# Turn annual drought values into raster 
raster_spei <- rasterFromXYZ(drought_201020)

# Make sure the CRS is consistent with shapefile
crs(raster_spei) <- "+proj=longlat +datum=WGS84 +no_defs"

# Calculate zonal statistics for each district 
bgd_spei_adm2 <- exact_extract(raster_spei, bgd_adm2, fun = c('mean'), append_cols=c("ADM2_PCODE", "ADM2_EN"))

# Remove "mean" from column names
names(bgd_spei_adm2) <- sub("^mean.", "", names(bgd_spei_adm2))

# Remove min SPEI value - we need a time series for 2010-2020
bgd_spei_adm2 <- bgd_spei_adm2[ , !names(bgd_spei_adm2) %in% 
    c("minSPEI","minSPEIzero")]

# Reshape long
long.drought <- melt(data.frame(bgd_spei_adm2), id.vars = c("ADM2_PCODE", "ADM2_EN"), variable.name = "year")

# Remove "SPEI" from the year values
long.drought$variable <- substr(long.drought$variable, 6, 10)

# Rename column as year
names(long.drought)[names(long.drought) == 'variable'] <- "year"

# Rename column to indicate SPEI values
names(long.drought)[names(long.drought) == 'value'] <- "SPEI"

# Create a variable assigning categories of drought
long.drought <- long.drought %>% mutate(Drought =
                     case_when(SPEI < -1.5 ~ "1", 
                               SPEI < -1 & SPEI > -1.5 ~ "2",
                               SPEI < -0.5 & SPEI > -1 ~ "3",
                               SPEI< 0 & SPEI > -0.5 ~ "4",
                               )
)

# Remove NA values
long.drought <-subset(long.drought, Drought!="NA")

# Count the number of states in each category of drought each year  
results2 <- long.drought %>%
  group_by(Drought, year) %>%
  summarize(count = n_distinct(ADM2_EN))

# Order by year and drought
results2 <- results2[order(results2$year, results2$Drought),]

# Expand a data frame to cover all categories of drought and all years 
g <- with(results2, expand.grid(year = seq(min(year), max(year)), Drought = unique(Drought), count = 0)) #

# Reorder the data frame
g <- g[order(g$year, g$Drought),]

# Change column name
colnames(results2)[3] <- "number"

# Turn years into numeric
g$year <- as.numeric(g$year)
results2$year <- as.numeric(results2$year)


# Left join - need for the results data frame to override the other one 
test <- left_join(g, results2)

# Replace values of count with number if number is non-missing
test$number <-replace(test$number, is.na(test$number), test$count)

# Drop the count column 
test = subset(test, select = -count )

# Rename codes
test$Drought[test$Drought == '1'] <- 'SPEI < -1.5'
test$Drought[test$Drought == '2'] <- '-1.5 < SPEI < -1'
test$Drought[test$Drought == '3'] <- '-1 < SPEI < -0.5'
test$Drought[test$Drought == '4'] <- "-0.5 < SPEI < 0"

# Give a specific order:
test$Drought <- factor(test$Drought , levels=c('SPEI < -1.5', '-1.5 < SPEI < -1', '-1 < SPEI < -0.5', "-0.5 < SPEI < 0") )


# Plot
ggplot(test, aes(x=year, y=number, fill=Drought)) + 
    geom_area()+ 
scale_x_continuous(breaks= pretty_breaks())+ theme_minimal() + scale_fill_brewer(palette="Blues")+
  xlab("Year") + ylab("Number of districts") + ggtitle("Number of districts affected by drought in Bangladesh in the period 2010-2020")

# Check with values from the drought data frame
summary(drought_201020)


```

Create a graph for number of floods per year during 2010-2018

```{r}
# Import csv file with flood data 
bihs_flood <- read.csv("C:/Users/idabr/OneDrive - Oxford Policy Management Limited/DEEP Conflict Climate Bangladesh/Data/Output/BIHS_flood.csv")

# Save column names from the flood data frame - those already contain the dates of the flood events
flood.dates <- colnames(bihs_flood[,7:45])

# Leave only the year of the flood
flood.dates <- substr(flood.dates, 3, 6) # remove day and month

# Turn into a table
flood.table <- table(flood.dates)
flood.table

# Turn into a data frame
flood.ts <- as.data.frame(flood.table)

# Drop levels
flood.ts$flood.dates <- levels(droplevels(flood.ts$flood.dates))

# Change to numeric
flood.ts$flood.dates <- as.numeric(flood.ts$flood.dates)
flood.ts$Freq <- as.numeric(flood.ts$Freq)

# Plot
ggplot(flood.ts, aes(x=flood.dates, y=Freq, fill="Blue", legend=F)) + 
    geom_area()+ 
scale_x_continuous(breaks= pretty_breaks())+ 
scale_y_continuous(breaks= pretty_breaks()) + 
  theme_minimal() + scale_fill_brewer(palette="Blues")+
  xlab("Year") + ylab("Number of floods") + ggtitle("Number of floods in Bangladesh in the period 2010-2018")


```


